Deterministic contention management for low latency Cloud RAN over an optical ring

### 2019-01-11 v3/articleV3.tex

In this article, we study a Cloud RAN (C-RAN) application based on the N-GREEN optical ring described in~\cite{ngreenarchitecture,uscumlic2018scalable}. C-RAN is one of the major area of development for 5G; it consists in centralizing or partially centralizing the computation units or {\bf BaseBand Units} (BBU) of the {\bf Remote Radio Heads} (RRH) in one datacenter~\cite{mobile2011c}. Periodically, each RRH on the field send some uplink traffic to its associated BBU in the datacenter, then, after a computation, the BBU send the same quantity of downlink traffic back to the RRH. The latency of the messages between the BBU and the RRH is critical since some services need end-to-end latency as low as $1$ms~\cite{3gpp5g,boccardi2014five}.

Nowadays, the traffic is managed by statistical multiplexing~\cite{kern2006applying}. Here, we propose an SDN approach to {\bf deterministically} manage the periodic C-RAN traffic by choosing emission timing. Indeed, Deterministic Networking is one of the strong current aspects considered to reduce the E2E latency~\cite{finn-detnet-architecture-08}. In a previous work~\cite{dominique2018deterministic}, the authors have studied a similar problem for a star shaped network. In contrast with our previous work, finding emission timings so that different periodic sources do not use the same resource is easy in the context of the N-GREEN optical ring. However, we add two new difficulties arising from practice: the messages from RRHs are scattered and there are other traffics whose latency must be preserved. It turns out that the deterministic management of CRAN traffic
we propose reduce the latency of CRAN traffic to the physical delay of the routes, while reducing the latency of the other traffics by smoothing the load of the ring in the period. However, to achieve such a good latency, our solution needs to reserve ressource in advance, which slightly decreases the maximal load the N-GREEN optical ring can manage. Such an approach of reservation of the network for an application (CRAN in our context) relates to network slicing~\cite{jiang2016network} or virtual-circuit-switched  connections in optical networks~\cite{cadere2010virtual,szymanski2016ultra}.

@@

   The RRHs are the source of the {\bf deterministic and periodic} C-RAN traffic.
   There are $k$ RRHs attached to the ring and several RRHs can be attached to the same vertex. An RRH is linked to a node of the ring through an electronic interface of bit rate $R$ Bps.
   The ring has a larger bit rate of $F\times R$ Bps. The integer $F$ is called the {\bf acceleration factor} between the electronic and the optical domains. A node aggregates the data received on the electronic interface during $F$ units of time to create a packet of size $C$ and then puts it in the insertion buffer.
%   fills a container of the ring with this data. 
  In each period $P$, an RRH emits data during a time called \textbf{emission time} or $ET$. Hence the RRH emits $ET / F$ packets, i.e. a container of size $C$ each $F$ units of time during the emission time, see Fig.~\ref{fig:interface}.
   %Each period $P$, an RRH emits $ET / F$ packets, i.e. a packet of size $C$ each $F$ units of time during a time $ET$ (emission time),

   The data of the RRH $i$ arrives at some node $u$ at a time $m_i$ called {\bf offset}. The offsets can be determined 
   by the designer of the system and can be different for each RRH but must remain the same over all periods. We assume that all BBUs are contained in the same data-center attached to the node $v$. The data from $u$ is routed to its BBU at node $v$ through the ring and arrives at time $m_i + \omega(u,v)$ if it has been inserted in the ring upon arrival. Then after some computation time (which w.l.o.g. is supposed to be zero), an answer is sent back from the BBU to the RRH. The same quantity of data is emitted by each BBU or RRH during any period.

@@

To avoid latency between the arrival of the data from the RRH and its insertion on the optical ring, 
we allow nodes to \textbf{reserve} a container one round before using it. A container which is reserved cannot be filled by any node except the one which has reserved it (but it may contain data when it is reserved). 
Let $u$ be the node to which is attached the RRH $i$, if $u$ reserves the container $u$ at time $m_i - RS$, then it is guaranteed that $u$ can fill a free container at time $m_i$ with the data of the RRH $i$.
In the method we now describe, the C-RAN packets never wait in the node: The message sent by the RRH $i$ arrives at its BBU at node $v$ at time $m_i + \omega(u,v)$ and the answer is sent from the BBU at time $m_i + \omega(u,v) +1$.

@@

Given a ring, a set of RRH's, a period and an acceleration factor $F$, the problem we solve here is to find an \textbf{assignment} of values of the $m_i$'s which is \textbf{valid}: two RRHs never reserve or use the same container in a period. Moreover we want to preserve the latency of the BE traffic. It means that the time a BE packet spent in the insertion buffer must be minimized. To do so, we must minimize the time a node waits for a free container at any point in the period, by spreading the C-RAN traffic as uniformly as possible over the period in order to give the nodes an available container in a minimal average time.

Fig.~\ref{fig:assignment} represent an assignment of three couples RRH-BBU. each slot has a duration $F$ unit of times, and, since an RRH/BBU emits a packet each $F$ UoT during $ET$ UoT, if we take the granularity of a slot to represent the time, the emission of a BBU/RRH is continuous in the time, during $ET$ UoT. Also, if we take a date $t$ in the period $P$, then $t$ is in the slot $\frac{P}{F}$ and is at position $P \mod F$.

@@

      \includegraphics[scale=0.25]{repart1res}
     \caption{Impact of compacting position on the BE traffic latency.}   \label{fig:algocmp}

>>>>

In this article, we study a Cloud RAN (C-RAN) application based on the N-GREEN optical ring described in~\cite{ngreenarchitecture,uscumlic2018scalable}. C-RAN is one of the major area of development for 5G; it consists in centralizing or partially centralizing the computation units or {\bf BaseBand Units} (BBU) of the {\bf Remote Radio Heads} (RRH) in one datacenter~\cite{mobile2011c}. Periodically, each RRH on the field sends some uplink traffic to its associated BBU in the datacenter, then, after a computation, the BBU sends the same quantity of downlink traffic back to the RRH. The latency of the messages between the BBU and the RRH is critical since some services need end-to-end latency as low as $1$ms~\cite{3gpp5g,boccardi2014five}.

Nowadays, the traffic is managed by statistical multiplexing~\cite{kern2006applying}. Here, we propose an SDN approach to {\bf deterministically} manage the periodic C-RAN traffic by choosing emission timing. Indeed, Deterministic Networking is one of the main method considered to reduce the E2E latency~\cite{finn-detnet-architecture-08}. In a previous work~\cite{dominique2018deterministic}, the authors have studied a similar problem for a star shaped network. In contrast with our previous work, finding emission timings so that different periodic sources do not use the same resource is easy in the context of the N-GREEN optical ring. However, we add two new difficulties arising from practice: the messages from RRHs are scattered and there are other traffics whose latency must be preserved. It turns out that the deterministic management of CRAN traffic
we propose reduce the latency of CRAN traffic to the physical delay of the routes, while reducing the latency of the other traffics by smoothing the load of the ring over the period. To achieve such a good latency, our solution needs to reserve ressources in advance, which slightly decreases the maximal load the N-GREEN optical ring can handle. Such an approach of reservation of the network for an application (CRAN in our context) relates to network slicing~\cite{jiang2016network} or virtual-circuit-switched connections in optical networks~\cite{cadere2010virtual,szymanski2016ultra}.

@@

   The RRHs are the source of the {\bf deterministic and periodic} C-RAN traffic.
   There are $k$ RRHs attached to the ring and several RRHs can be attached to the same vertex. An RRH is linked to a node of the ring through an electronic interface of bit rate $R$ Bps.
   The ring has a larger bit rate of $F\times R$ Bps. The integer $F$ is called the {\bf acceleration factor} between the electronic and the optical domains. A node aggregates the data received on the electronic interface during $F$ units of time to create a packet of size $C$ and then puts it in the insertion buffer.
%   fills a container of the ring with this data. 
  In each period $P$, an RRH emits data during a time called \textbf{emission time} or $ET$. Hence the RRH emits $ET / F$ packets, i.e. a container of size $C$ each $F$ units of time during the emission time, as shown in Fig.~\ref{fig:interface}.
   %Each period $P$, an RRH emits $ET / F$ packets, i.e. a packet of size $C$ each $F$ units of time during a time $ET$ (emission time),

   The data of the RRH $i$ arrives at some node $u$ at a time $m_i$ in a period called {\bf offset}. The offsets can be determined by the designer of the system and can be different for each RRH but must remain the same over all periods. We assume that all BBUs are contained in the same data-center attached to the node $v$. The data from $u$ is routed to its BBU at node $v$ through the ring and arrives at time $m_i + \omega(u,v)$ if it has been inserted in the ring upon arrival. Then after some computation time (which w.l.o.g. is supposed to be zero), an answer is sent back from the BBU to the RRH. The same quantity of data is emitted by each BBU or RRH during any period.

@@

To avoid latency between the arrival of the data from the RRH and its insertion on the optical ring, 
we allow nodes to \textbf{reserve} a container one round before using it. A container which is reserved cannot be filled by any node except the one which has reserved it (but it may not be free when it is reserved). 
If $u$ reserves a container at time $m_i - RS$, then it is guaranteed that $u$ can fill a free container at time $m_i$ with the data of the RRH $i$.
In the method we now describe, the C-RAN packets never wait in the node: The message sent by the RRH $i$ arrives at its BBU at node $v$ at time $m_i + \omega(u,v)$ and the answer is sent from the BBU at time $m_i + \omega(u,v) +1$.

@@

Given a ring, a set of RRH's, a period and an acceleration factor $F$, the problem we solve here is to find an \textbf{assignment} of values of the offsets $m_i$'s which is \textbf{valid}: two RRHs never use the same container in a period. Moreover we want to preserve the latency of the BE traffic. It means that the time a BE packet waits in the insertion buffer must be minimized. To do so, we must minimize the time a node waits for a free container at any point in the period, by spreading the C-RAN traffic as uniformly as possible over the period in order to give the nodes an available container in a minimal average time.

Figure~\ref{fig:assignment} represents an assignment of three couples of RRH and BBU. Each slot has a duration of $F$ unit of times, and, since an RRH/BBU emits a packet each $F$ UoT during $ET$ UoT, if we take the granularity of a slot to represent the time, the emission of a BBU/RRH is continuous in our representation, during $ET/F$ slots. A date $t$ in the period corresponds in Figure~\ref{fig:assignment} to the slot $t/F$ and is at position $t \mod F$.

@@

      \includegraphics[scale=0.25]{repart1res}
     \caption{Impact of compacting position on the BE traffic latency.}   \label{fig:algocmp}
\end{center}
  \end{figure}
